%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
% Include in your acknowledgements any person who has made a substantial contribution to your research, but doesn't qualify as an author. Examples include other students, staff, projects or organisations which have provided financial support or data.
% If you have questions about this please get in touch with your supervisor!


\begin{document}

    \begin{acknowledgements}
      % If you have filled out \phdproject and \phdsupervisor in report.tex, you should use the \phdack to acknowledge the PhD project in your research.
        % \phdack\\
        
        I would like to thank my supervisors Prof. Dr. Sebastian Houben and Prof. Dr.-Ing. Udo Frese for their guidance and support throughout my research. I express my gratitude to M. Sc. Raúl Domínguez and Dr.-Ing. Christoph Hertzberg for their valuable insights and feedback on my work. This thesis work would not have been possible without their support and encouragement. 
        
        I would like to thank my colleagues at \gls{dfki} \gls{ric} for their support and collaboration during my research. Special thanks to Felix Glinka, Mihaela Popescu and Dr. Ing. Christoph Hertzberg for their time and effort in helping me understand the concepts and techniques used in this research. 
        
        I would also like to thank Raúl Domínguez, Christoph Hertzberg and Kabilan Tamilmani for their support during the experiments and data collection phase of this research. I would like to thank the \gls{dfki} \gls{ric} for providing the opportunity and the resources to conduct this research. 
        
        A special thanks to my friends, Oviya Rajavel, Kabilan Tamilmani, Tharun Sethuraman, Selvakumar Nachimuthu, and Akhilan Ashokan for their support and encouragement which helped me to stay motivated and focused throughout my research. 
        
        Finally, I would like to thank my family for their unwavering support and encouragement throughout my masters program. Their love and support have been invaluable to me, and I am grateful for their presence in my life.
       
    \end{acknowledgements}
\end{document}
