%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}

    \chapter{Additional Implementation Details}
    
    This chapter provides additional implementation details of the \gls{cvo} approach. It includes the derivation of the measurement Jacobian, the numerical Jacobian calculation, and a comparison of contact estimation methods. 
    
    \section{Measurement Jacobian Derivation} \label{appendix:MeasurementJacobianDerivation}
    
    This section provides the derivation of the measurement Jacobian presented in Section~\ref{sec:measurementModelAndJacobian}. The Jacobian is derived by calculating the partial derivatives of the measurement function $h(x)$ with respect to each error-state variable. This analytical derivation used to get the Jacobian matrix is derived with AI assistance using Gemini CLI~\cite{geminicli2026}. The presented equations in the following sections are generated using the AI assistant ChatGPT~\cite{openai_chatgpt_jacobian_2026}, and they have been manually verified for correctness. 
    
    \subsection*{Rotation Matrix}

    The rotation matrix block of the measurement Jacobian is derived from the measurement function $h(x)$, which is defined as follows:

    \begin{equation}
    h(x) =
    -\left(
    [R_{body}^{odom}]^{T}v_{body}^{odom}
    +
    \left(\omega_{body}^{body}\times r_{pc}^{body}\right)
    \right)
    \end{equation}

    For a small rotation error $\delta R$, the first-order approximation of the body velocity is
        
    \begin{equation}
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        \approx
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        +
        \left[
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        \right]_{\times}\delta R
    \end{equation}

    Substituting this approximation into the measurement function $h(x)$, we get,

    \begin{align}
        h(x,\delta R)
        &\approx
        -\left(
        [R_{body}^{odom}]^{T}v_{body}^{odom} + \left[ [R_{body}^{odom}]^{T}v_{body}^{odom} \right]_{\times}\delta R +
        \left(
        \omega_{body}^{body}\times r_{pc}^{body}
        \right)
        \right)
        \\
        &=
        -\left(
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        +
        \left(
        \omega_{body}^{body}\times r_{pc}^{body}
        \right)
        \right)
        -
        \left[
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        \right]_{\times}\delta R
        \\
        &=
        h(x)
        -
        \left[
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        \right]_{\times}\delta R.
    \end{align}
        
    Therefore, the partial derivative of the measurement function with respect to the rotation error is
    
    \begin{equation}
    \frac{\partial h(x,\delta R)}
    {\partial \delta R}
    \bigg|_{\delta R=0}
    =
    -\left[
    [R_{body}^{odom}]^{T}v_{body}^{odom}
    \right]_{\times}.
    \end{equation} 
    
    \subsection*{Velocity}

    A small velocity error $\delta V$ changes the measurement function as

    \begin{equation}
        h(x,\delta V)
        =
        -\left(
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        +
        \delta V
        +
        \left(
        \omega_{body}^{body}\times r_{pc}^{body}
        \right)
        \right).
    \end{equation}

    Using the definition of $h(x)$, this can be written as

    \begin{equation}
        h(x,\delta V)
        =
        h(x)-\delta V.
    \end{equation}

    Differentiating with respect to $\delta V$ gives

    \begin{equation}
        \frac{\partial h(x,\delta V)}
        {\partial \delta V}\bigg|_{\delta V=0}
        =
        -\mathbf{I}.
    \end{equation}

    \subsection*{Position}

    The body position does not appear in the measurement function. Therefore, adding a small position error $\delta P$ does not change the predicted measurement:

    \begin{equation}
        h(x,\delta P)=h(x).
    \end{equation}

    Differentiating with respect to $\delta P$ gives

    \begin{equation}
        \frac{\partial h(x,\delta P)}
        {\partial \delta P}\bigg|_{\delta P=0}
        =
        \mathbf{0}.
    \end{equation}

    \subsection*{Gyroscope Bias}

    The gyroscope bias is included in the bias-corrected angular velocity:

    \begin{equation}
        \omega_{body}^{body}
        =
        \omega_{meas}^{body}-b_g.
    \end{equation}

    After adding a small gyroscope-bias error $\delta b_g$, the measurement function becomes

    \begin{align}
        h(x,\delta b_g)
        &=
        -\left(
        [R_{body}^{odom}]^{T}v_{body}^{odom}
        +
        \left(
        \omega_{meas}^{body}
        -
        b_g
        -
        \delta b_g
        \right)
        \times r_{pc}^{body}
        \right)
        \\
        &=
        h(x)
        +
        \delta b_g\times r_{pc}^{body}
        \\
        &=
        h(x)
        -
        [r_{pc}^{body}]_{\times}\delta b_g.
    \end{align}

    Differentiating with respect to $\delta b_g$ gives

    \begin{equation}
        \frac{\partial h(x,\delta b_g)}
        {\partial \delta b_g}\bigg|_{\delta b_g=0}
        =
        -[r_{pc}^{body}]_{\times}.
    \end{equation}

    \subsection*{Accelerometer Bias}

    The accelerometer bias does not appear directly in the measurement function. Therefore, adding a small accelerometer-bias error $\delta b_a$ does not change the predicted measurement:

    \begin{equation}
        h(x,\delta b_a)=h(x).
    \end{equation}

    Differentiating with respect to $\delta b_a$ gives

    \begin{equation}
        \frac{\partial h(x,\delta b_a)}
        {\partial \delta b_a}\bigg|_{\delta b_a=0}
        =
        \mathbf{0}.
    \end{equation}

    \section{Numerical Jacobian} \label{appendix:NumericalJacobianCalculation}

    The analytical measurement Jacobian derived in Appendix~\ref{appendix:MeasurementJacobianDerivation} is verified using a numerical Jacobian. Each error-state variable is perturbed independently, and the corresponding change in the measurement function is calculated using the central difference approximation. This numerical Jacobian validation implementation and the following equations were developed with the help of AI assistant ChatGPT~\cite{openai_chatgpt_jacobian_2026}. The script is carefully reviewed and verified for correctness. 

    For the $i$-th error-state variable, the corresponding column of the numerical
    Jacobian is calculated as

    \begin{equation}
        H_{\mathrm{num},i}
        =
        \frac{
            h\left(x\boxplus\epsilon_i e_i\right)
            -
            h\left(x\boxplus(-\epsilon_i e_i)\right)
        }{2\epsilon_i},
        \label{eq:numericalJacobianCentralDifference}
    \end{equation}

    where, $e_i$ is the unit vector corresponding to the $i$-th error-state variable, $\epsilon_i$ is the applied perturbation, and $\boxplus$ represents the addition of the perturbation to the state. The rotation perturbation is applied on $SO(3)$ using

    \begin{equation}
        R^{+}
        =
        R\exp\left([\delta R]_{\times}\right),
    \end{equation}

    while the velocity, position, and bias perturbations are added directly to their corresponding state variables. For the selected test state and perturbation values, the analytical and numerical Jacobians were consistent. The Python implementation used for this verification is provided in Listing~\ref{lst:numericalJacobianVerification}.

    \lstinputlisting[
        language=Python,
        caption={Numerical verification of the measurement Jacobian using central differences.},
        label={lst:numericalJacobianVerification},
        frame=single,
    ]{media/config/numerical_jacobian.py}

    The analytical and numerical Jacobians obtained for the selected test state are shown in Listing~\ref{lst:numericalJacobianOutput}. Both Jacobians contain the same values, which verifies the analytical derivation for this test case.

    
    \begin{lstlisting}[
        caption={Comparison of the analytical and numerical measurement Jacobians.},
        label={lst:numericalJacobianOutput},
        basicstyle=\ttfamily\footnotesize
    ]
    
    Expected Jacobian:

    0 -0.2 -0.5 -1.0    0    0    0    0    0    0  0.8 -0.4    0    0    0
    0.2    0  1.0    0 -1.0    0    0    0    0 -0.8    0  0.5    0    0    0
    0.5 -1.0    0    0    0 -1.0    0    0    0  0.4 -0.5    0    0    0    0

    Computed numerical Jacobian:

    0 -0.2 -0.5 -1.0    0    0    0    0    0    0  0.8 -0.4    0    0    0
    0.2    0  1.0    0 -1.0    0    0    0    0 -0.8    0  0.5    0    0    0
    0.5 -1.0    0    0    0 -1.0    0    0    0  0.4 -0.5    0    0    0    0
    \end{lstlisting}

    \section{Contact Estimation Comparison} \label{sec:contactEstimationComparison}
    
    \todo[inline]{Provide the contact estimation comparison mahalanobis distance and lowest foot contact}
        
    \chapter{Parameters}

    This chapter provides the OroGen and YAML configuration files of the \gls{cvoe} component. These are the configurations used in the experiments and evaluations presented in this thesis. The OroGen file defines the component's structure, including its input and output ports, operations, and properties. The YAML configuration file specifies the parameters for the component, such as filter settings, hardware configurations, and frame names. The OroGen file is provided in Listing~\ref{code:orogenfileCVOE}, while the YAML configuration file is presented in Listing~\ref{code:YAMLConfigfileCVOE}. Additionally, a sample summary of the statistical error metrics obtained from the evaluation of the \gls{cvoe} component is provided in Listing~\ref{lst:statistical-error-metrics}.
    
    \lstinputlisting[style=orogenstyle,
    caption={OroGen file of the \gls{cvoe} component},
    label={code:orogenfileCVOE}]{media/config/contact_velocity_odometry_estimator.orogen}

    \lstinputlisting[style=yamlstyle,
    caption={Contact Velocity Odometry Estimator YAML configuration file},
    label={code:YAMLConfigfileCVOE}]{media/config/contact_velocity_odometry_estimator::Task.yml}
        
    \begin{lstlisting}[style=consoleoutput, caption={Sample summary of the statistical error metrics.}, label={lst:statistical-error-metrics}]
    
    --------------------------------------------------
    Position error metrics
    --------------------------------------------------
    Max position error          : 0.92 m
    Mean position error         : 0.32 m
    Min position error          : 0.00 m
    Std dev of position error   : 0.32 m
    RMSE of position error      : 0.45 m    
    --------------------------------------------------
    Orientation Geodesic angle error metrics
    --------------------------------------------------
    Max geodesic angle error        : 9.36 degrees
    Mean geodesic angle error       : 4.80 degrees
    Min geodesic angle error        : 0.00 degrees
    Std dev of geodesic angle error : 2.65 degrees
    RMSE of geodesic angle error    : 5.49 degrees
    --------------------------------------------------
    --------------------------------------------------
    Relative Pose Error (RPE) Translation metrics
    --------------------------------------------------
    Max translation error                   : 0.05 m
    Mean translation error                  : 0.04 m
    Std dev of translation error            : 0.00 m
    RMSE of translation error               : 0.04 m

    Mean translation error per meter        : 0.15 m/m
    Std dev of translation error per meter  : 0.01 m/m
    RMSE of translation error per meter     : 0.15 m/m
    --------------------------------------------------
    Relative Pose Error (RPE) Rotation metrics
    --------------------------------------------------
    Max rotation error                  : 3.59 degrees
    Mean rotation error                 : 0.91 degrees
    Std dev of rotation error           : 0.92 degrees
    RMSE of rotation error              : 1.29 degrees

    Mean rotation error per meter       : 3.62 degrees/m
    Std dev of rotation error per meter : 3.70 degrees/m
    RMSE of rotation error per meter    : 5.18 degrees/m
    --------------------------------------------------
    \end{lstlisting}
    
    \chapter{[in the case of using AI-assistants] Description of AI-Generated Content}
    Describe in detail what content was generated using an AI assistant.
    Name the AI assistant and how it was used (e.g. which prompts were used, and for which parts of the project).
    
    ChatGPT plus was used. 
    
    - the UML diagrams, 
    - grammar and sentence structure corrections,
    - latex table from excel data,
    - latex code for equations from a rudimentary description of the equations,
    - coherence and flow checking of the report
    Summarizing overview of results and conclusion from the results and discussion chapter.
\end{document}
