%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}
    \chapter{Introduction} \label{chap:introduction}

    For any mobile robot, a reliable estimate of its pose relative to its environment is crucial for autonomous navigation. In general, mobile robots estimate their pose based on the incremental motion calculated using proprioceptive sensors such as \gls{imu} and wheel encoders. This pose estimation technique is called dead-reckoning based odometry. In robotic platforms with limited onboard computational power, this method of odometry is particularly important due to its simplicity and computational efficiency \cite{Agostinho2022}. This approach is vital in scenarios where positioning systems like \gls{gps}/\gls{gnss} and Motion capture systems are unavailable. The odometry estimate based on a single source of odometry is susceptible to sensor drift over time and wheel slip related errors.
    
    While these errors are prevalent across mobile robots, their effects are particularly evident in the systems operating in slip-prone terrain. One such system is the Coyote~3 rover. This rover uses a non-traditional wheel type, the \gls{rw}, which is a hybrid between wheeled and legged locomotion. The spokes provide a foot-like terrain interaction while the overall wheel structure keeps the motion and control simple. These rovers, due to their spoked wheels, can traverse diverse terrains like soft and coarse sand terrain, rocky terrain and inclined terrain. Unlike traditional wheeled rovers, rimless wheeled rovers can overcome challenging obstacles effectively \cite{Brinkmann2018}.
    
    This hybrid locomotion capability of the \gls{rwr} system comes at a cost. The spokes of the \gls{rwr} system change the wheel-ground interaction from smooth rolling motion to discontinuous contact-based motion. The current baseline odometry in the Coyote~3 rover assumes a circular wheel model and uses wheel odometry for translational estimation and \gls{imu} data for orientation estimation. Since the translation and rotation are decoupled, this is commonly referred to as decoupled odometry. This approach provides a simple means of odometry estimation without complex state estimation techniques. Any slippage that occurs when traversing slip-prone terrain is not accounted for and the pose error from the slippage propagates to the odometry estimate. To address this problem, sensor fusion techniques are often employed.
    
    This project focuses on improving the proprioceptive odometry using sensor fusion techniques, specifically the \gls{inekf}. We also focus on slip detection and excluding wheel measurements affected by slippage. Finally, a framework that estimates the spokes in contact has also been developed. The remainder of this chapter is organized as the following subsections: the motivation, the current challenges for this project, and an outline of this project report.

    \section{Motivation}
    
    In harsh environment conditions like underground cave exploration, search and rescue missions in a collapse-prone building and space exploration mission, the \gls{rwr} is a suitable candidate. Such conditions often deny the possibility of external state estimation methods like \gls{gps} or Motion Capture Systems. In such scenarios, the importance of a reliable pose estimation based on proprioceptive sensors is paramount.
    
    The current baseline odometry in \gls{rwr} does not explicitly account for slip-related error in such harsh environments. The rotation of the wheels during slippage is also considered robot motion. This flaw makes this odometry less reliable in slippery terrain like loose sand and inclined surfaces. Baseline odometry also ignores the hybrid nature of the \gls{rw} and assumes a circular wheel model with an effective wheel radius. This assumption fails to include the varying geometry of the \gls{rw} and the dynamics of making and breaking contacts. The use of only \gls{imu} data causes sensor drift in the orientation estimation and the use of Wheel encoder data causes the slip related errors to be included in the translation estimation in the baseline odometry. Hence, a reliable slip-aware odometry for the \gls{rwr} in varied terrain conditions remains a challenging problem. 
    
    \section{Contributions}
    
    This thesis aims to improve the pose estimation for the Coyote~3 \gls{rwr} system. An \gls{inekf} based sensor fusion framework is developed to fuse the \gls{imu} sensor data and the wheel encoder data to estimate the pose of the rover. This framework estimates the rover state using \gls{imu} data and corrects this estimation by using the wheel measurements. When inconsistency between these two sources of information occurs, the filter rejects the wheel measurements based on the filter's innovation uncertainty, thereby rejecting the measurement update. The contact state of the wheel spokes can be probabilistically inferred from the innovation uncertainty.
        
    \section{Report Outline}

    The report structure is as follows:
    
    \begin{itemize}
        \item[] Chapter \ref{chap:background} presents the background details on the Coyote~3 \gls{rwr} system, the \gls{inekf}, wheel slip, the notations and conventions used in this report.
        \item[] Chapter \ref{chap:SOTA} provides information on previous works related to proprioceptive odometry, \gls{inekf} based sensor fusion techniques and other relevant odometry estimation techniques and their limitations.
        \item[] Chapter \ref{chap:problem_statement} focuses on the problem in Coyote~3 odometry system, limitations of baseline odometry and challenges in solving this problem.
        \item[] Chapter \ref{chap:methodology} describes the proposed slip-aware odometry using contact point velocity, the Mahalanobis-distance-based measurement rejection and the contact estimation.
        \item[] Chapter \ref{chap:solution} presents the details on Coyote~3 system implementation, coordinate frame convention, assumptions used and challenges in using real time sensor data.
        \item[] Chapter \ref{chap:experimentandevaluation} provides details on the experimental setup, collected experimental datasets, the evaluation metrics used, the evaluation of proposed approach against baseline odometry and ground truth data.
        \item[] Chapter \ref{chap:resultsanddiscussion} discusses the evaluation results and the performance of the proposed approach against the ground truth and baseline odometry in general and edge-case scenarios.
        \item[] Chapter \ref{chap:conclusionandfuturework} provides summarized results, findings and limitations of the proposed approach and discusses the potential future improvements in the proposed approach.
    
    \end{itemize}

\end{document}
