%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}    
\chapter{State of the Art}
\label{chap:sota}

This chapter provides details on the relevant state-of-the-art methods for odometry and slip estimation in legged and wheeled robots. Since the prevalence of \gls{rwr} is low in the literature compared to legged and wheeled robots, this chapter summarizes the methods used in legged and wheeled robots, which can be adapted to \gls{rwr}. This chapter is organized into the following sections: Wheel Inertial Odometry Fusion, Contact-aware state estimation, Slip detection and rejection, Learning-based slip estimation, and research gap. 

\section{Wheel Inertial Odometry Fusion}
\label{sec:sota_wheel_inertial}

    The most common proprioceptive sensors used for odometry estimation are wheel encoders and \gls{imu}. Wheel encoders measure wheel rotation, which can be converted into an estimate of the robot’s motion using the wheel kinematic model. The \gls{imu} provides information about the robot's angular velocity and linear acceleration, which can be used to estimate the robot's orientation and the acceleration can be integrated to estimate the robot's velocity and position. 
    
    However, wheel encoders are prone to errors due to wheel slippage, which can occur on inclined and slippery terrain where the wheel slippage is more prominent. The \gls{imu} is prone to accumulated error from the integration of the angular velocity and linear acceleration, which can lead to drift in the estimated position and orientation over time. To combat these limitations, wheel encoder and \gls{imu} data are often fused together to get a reliable estimate of the robot's state. This fusion process is commonly done using a Kalman filter, which is an algorithm that uses noisy sensor measurements to estimate a system's state. Some variants of the Kalman Filter, such as the \gls{ekf}, Error State Kalman Filter (ESKF), and Unscented Kalman Filter (UKF) are used for nonlinear systems across the literature. 
    
    Conventional wheel inertial fusion approaches fuse wheel encoder and \gls{imu} data to estimate the robot's state. This reduces the weakness of individual sensors and provides a more accurate estimate of the robot's state. This combination of complementary sensors can provide a more reliable state estimation \cite{Yousuf2016}. The estimated state is later combined with \gls{gps} data. This has been shown to improve the convergence and accuracy of state estimation. Although this approach improves the position estimate, it does not explicitly handle wheel slip or estimate the robot's orientation.
    
    Unlike conventional fusion approaches that use a fixed measurement covariance, robust filtering methods reduce the influence of wheel odometry measurements that are inconsistent with the predicted state. Such methods use variations of the Kalman filter, such as the Huber regression Kalman filter, the covariance scaling Kalman filter and the variational Bayesian filters. In the covariance scaling Kalman filter, the ratio of \gls{md} and the chi-square distribution is used to adapt the covariance of the measurements when abnormal wheel measurements are detected \cite{Das2021}. The robust-filter variants generally reduced the 3D position \gls{rmse} compared with the standard wheel inertial filter, although the best-performing method differed between datasets. However, these approaches are tested on a rover with conventional wheels, which provide a stable contact with the ground, and they do not represent the contact dynamics of the \gls{rw}. 
    
    While robust filtering methods reduce the influence of wheel odometry measurements, model-based approaches attempt to estimate the slip through kinematic modelling and compensate for it. One such approach, a multi-innovation UKF, uses a sequence of current and previous innovations during the correction step to improve the estimation accuracy. In this method, the robot's state is initially estimated using encoder data processed through Instantaneous Center of Rotation (ICR) kinematics, which is then fused with inertial odometry for improved accuracy. However, a limitation of this specific method is its reliance on the ICR kinematics assumption that wheels on the same side have identical angular velocities. This assumption may not hold true under varying slip conditions \cite{Liu2020}. In addition, the assumption of continuous ground contact is not valid for \glspl{rw}. 
    
    Overall, wheel inertial fusion methods differ mainly in the way they handle wheel slip. Conventional fusion methods do not explicitly handle wheel slip, whereas robust filtering methods reduce the impact of inconsistent wheel-odometry measurements. Model-based approaches attempt to estimate the slip through kinematic modelling and compensate for it. However, all these approaches are based on a rover with conventional wheels, which provide a continuous contact with the ground, and they do not represent the contact dynamics of the \gls{rw}. Hence, contact-aware state estimation methods are explored in the next section.
    
\section{Contact-aware State Estimation}
\label{sec:sota_contact_aided}

    Contact-aided state-estimation methods use information about the interaction between the robot and the terrain to constrain the estimated motion. In these approaches, the robot state is propagated using \gls{imu} measurements, while the position or velocity of a point in contact with the ground is used during the correction step. When a contact point remains stationary, it provides an additional constraint that can reduce the drift caused by the integration of \gls{imu} measurements. Some contact-aided filtering approaches also augment the filter state with the positions of the active contact points \cite{Li2023,Kim2025}. However, the accuracy of these methods depends on correctly detecting the contact state and determining whether the contact point is stationary. If a slipping contact is incorrectly treated as stationary, the contact constraint can introduce an error into the state estimate.
        
    A contact-aided \gls{inekf} was developed for legged-robot state estimation. The filter propagates the robot pose, velocity, and \gls{imu} biases using inertial measurements, while forward kinematics of the contacting feet are used during the correction step. The positions of the active contact points are included in the filter state and are added or removed when the contact state changes. By defining the estimation error using the Lie-group structure of the state, the linearised error dynamics and observation model become independent of the current state estimate. The approach was evaluated in simulation and on a biped robot, where it showed improved convergence and consistency compared to a conventional quaternion-based \gls{ekf} \cite{Hartley2018ContactAidedInEKF}.

    Contact-aided invariant filtering has mainly been applied to legged systems. One such application uses a right-invariant \gls{ekf} to integrate \gls{imu} measurements, joint kinematics, foot-force measurements, and centre-of-mass dynamics for a biped robot \cite{Li2023}. The position of the supporting foot is included in the filter state and is used as a contact constraint. Since this constraint is affected when the foot slips, centre-of-mass dynamics are added as an additional source of information that is less sensitive to contact slip. The authors report that this reduced the floating-base velocity error in the direction affected by slip. However, the approach requires foot force measurements and a dynamic model of the robot, and it does not explicitly reject measurements of slipping contact.

    For wheeled-legged robots, the contact point cannot always be treated as a fixed point because the wheel contact moves along the wheel surface during rolling. The authors address this problem by modelling the moving wheel ground contact point together with wheel-induced displacement and probabilistic slip \cite{Zhou2025}. Their method combines \gls{imu} measurements, motor encoders, joint torque measurements, and wheel leg kinematics in a Kalman filter-based estimator. The results show that modelling the moving contact point improves the position and velocity estimation across wheeled, legged, and hybrid locomotion modes. However, the method assumes a conventional wheel with a single contact point moving continuously along the surface. This assumption does not directly represent a \gls{rw}, where contact changes between separate spokes.

    Contact-state uncertainty in wheel-legged systems has also been addressed by combining several contact indicators. The authors fuse gait phase, joint torque, and wheel-bottom height to estimate the probability of wheel ground contact \cite{Zhang2025}. The estimated probability is converted into a binary contact state and used together with encoder and \gls{imu} measurements for position and velocity estimation. This approach reduces the dependence on a single contact indicator, but the contact decision is based on a manually selected threshold, and the method is evaluated only in simulation. In addition, it does not explicitly estimate or reject slip-affected measurements.

    Instead of relying only on a binary contact decision, adaptive contact-aided filters can change the uncertainty assigned to the contact model. Kim et al. \cite{Kim2025} propose an adaptive \gls{inekf} that estimates the contact-model noise separately for each direction. Small foot motions are handled by increasing the corresponding contact uncertainty, while larger slip events can be rejected using the Mahalanobis distance thresholds. The results show that combining adaptive contact uncertainty with slip rejection improves the velocity and orientation estimation during dynamic motion on rough terrain. However, the method assumes legged contacts and augments the positions of the contacting feet in the filter state. It does not address the selection of a contact among multiple rimless-wheel spokes or use the velocity of the spoke tip as a measurement.

    Overall, contact-aided state-estimation methods use terrain contacts to constrain inertial state propagation and reduce drift. Their performance depends on the contact model, the reliability of contact detection, and the validity of the stationary-contact assumption. Methods developed for legged robots generally assume fixed foot contacts, while wheel-legged approaches model a single contact point moving continuously along the wheel surface. Neither assumption directly represents the intermittent and geometry-dependent contacts of a \gls{rw}. Methods for detecting slip and reducing the influence of unreliable contact or wheel measurements are discussed in the following section.


\section{Slip Detection and Slip-Aware Odometry}
\label{sec:sota_slip}

    Wheel slip causes a difference between the predicted motion from the wheel rotation and the actual motion of the robot. Slip-aware odometry methods attempt to reduce the effect of this difference on the state estimate. The previous works handle slip in different ways. Some methods estimate slip using a vehicle or terrain model and compensate for its effect, while others reduce the confidence in wheel measurements or reject them when slip is detected. This section reviews state-of-the-art methods for slip detection and for odometry that reject or adapt to the influence of slip-affected measurements. The methods are grouped into three categories: model-based slip estimation, innovation-based detection and measurement rejection, and other proprioceptive slip-detection methods.

    \subsection{Model-Based Slip Estimation}
    \label{subsec:sota_model_slip}

    Model-based slip-estimation methods use the relationships among wheel motion, robot motion, and terrain interaction to estimate slip. A dynamic vehicle model was combined with an \gls{ekf} to estimate the linear velocity, wheel velocities, sensor biases, and external disturbance acting on a conventional wheeled robot. The estimated longitudinal slip was then compared with a threshold to detect complete immobilization. The method detected immobilization using only proprioceptive measurements even when \gls{gps} information was unavailable. However, its accuracy depends on terrain and rolling-resistance parameters, and the model assumes continuous contact between the wheels and the terrain \cite{Ward2008}.

    A more detailed dynamic model can also be used to estimate longitudinal and lateral slip and correct wheel odometry measurements. One such approach models the interaction between compliant wheels and deformable soil using terramechanics equations. The slip estimated by the dynamic model is used to correct the rover velocity before the pose is propagated. This improves the dead-reckoning estimate compared with uncorrected wheel odometry. However, the method requires prior information about the rover, terrain, and soil properties. The evaluation also assumes that the soil properties remain homogeneous along the trajectory \cite{Gargiulo2021}.

    Model-based methods can explicitly estimate and compensate for slip, but their accuracy depends on the assumptions made in the vehicle and terrain models. Most of these methods assume conventional circular wheels with continuous wheel ground contact. These assumptions do not directly represent the intermittent contacts of a \gls{rw}, where the active contact changes between individual spokes.

    \subsection{Innovation-Based Detection and Measurement Rejection}
    \label{subsec:sota_innovation_slip}

    Instead of using a detailed terrain model, filter-based methods can detect unreliable measurements by comparing a measurement with its prediction from the current state estimate. A large innovation indicates that the measurement is not consistent with the predicted state. Robust wheel inertial filters use this information to reduce the influence of inconsistent wheel-odometry measurements. For example, the measurement covariance can be increased when the squared \gls{md} exceeds a chi-squared threshold. This reduces the effect of outliers without estimating a physical slip ratio, although the best-performing robust-filter method can vary between datasets \cite{Das2021}.

    Slip detection can also be followed by complete rejection of the wheel-odometry measurement. A slip ratio calculated from the wheel angular velocity, wheel radius, and estimated body velocity was used to detect severe slip in an automated guided vehicle. When the slip ratio exceeded a predefined threshold, wheel odometry was discarded, and the estimator relied on the \gls{imu} measurements. This reduced the error caused by wheel slip in the reported indoor experiment. However, the slip-ratio model assumes a fixed wheel radius and continuous rolling contact \cite{Xiaobo2024}.

    Contact measurements can also be handled by increasing their uncertainty instead of immediately rejecting them. An invariant fixed-lag smoother for a legged robot detects slip using the estimated foot velocity. When the velocity exceeds a threshold, the covariance of the stationary-contact constraint is increased. The smoother can also re-evaluate the contact state and previous foot velocities within its estimation window. This allows temporary contact errors to be corrected after additional measurements become available. However, the method assumes articulated foot contacts and does not model wheel-encoder-based contact velocity \cite{Yoon2024}.

    Adaptive contact-aided filtering combines covariance adaptation with measurement rejection. Direction-dependent contact uncertainty can be estimated and used to reduce the influence of small contact motions, while larger slip events can be rejected using the \gls{md}. This allows the estimator to distinguish between small violations of the stationary-contact assumption and larger slipping events. However, the approach is developed for legged robots and uses the positions of contacting feet in the filter state. It does not address the selection of contacts among several possible rimless-wheel spokes \cite{Kim2025}.

    These approaches avoid the need for a detailed terrain model and can reduce the influence of unreliable measurements. Their performance, however, depends on the predicted measurement, the assigned process and measurement covariances, and the threshold used to identify an inconsistent measurement.

    \subsection{Other Proprioceptive Slip-Detection Methods}
    \label{subsec:sota_other_slip}

    Slip can also be estimated using additional proprioceptive measurements related to wheel-terrain interaction. A slip-aware estimator for a wheeled-legged robot calculates a slip ratio using wheel motion, motor torque, and ground-contact information. The estimated slip ratio is then used to adapt the covariance of the wheel-derived measurements. When a severe slip is detected, the wheel measurements are assigned little influence, and the estimator mainly relies on the \gls{imu}. This approach improves position and velocity estimation during acceleration, braking, and intermittent wheel contact. However, it requires ground-pressure information and an effective-wheel-radius model and is limited to flat terrain in its current form \cite{Xu2024}.

    Another proprioceptive approach uses a network of local unscented Kalman filters (UKFs) to estimate the body velocity and longitudinal slip of a six-wheel planetary rover. Measurements from the \gls{imu}, wheel encoders, and steering joints are fused using a track-to-track fusion method. This allows the longitudinal slip ratio of different wheel sets to be estimated without an external positioning system. However, the approach assumes negligible lateral slip and continuous contact of fixed-radius wheels, and it has only been evaluated in simulation \cite{Zarei2024}.

    Overall, slip-aware odometry methods either explicitly model slip, reduce the influence of inconsistent measurements, reject slip-affected measurements, or adapt the measurement covariance based on an estimated slip condition. Model-based approaches can provide a direct slip estimate but require accurate vehicle and terrain models. Measurement-rejection and covariance-adaptation methods require fewer terrain parameters but depend on the consistency of the filter model and its uncertainty. The reviewed methods primarily estimate slip for complete wheels, wheel sets, or fixed-foot contacts. They do not directly detect slip at individual potential contact points of a \gls{rw}. Learning-based approaches that attempt to model slip from recorded sensor data are discussed in the following section.

    \section{Learning-Based Slip Estimation}
    \label{sec:sota_learning_slip}

    Learning-based methods have been investigated for slip detection and for improving odometry. One approach uses a Support Vector Machine (SVM) classifier to identify whether a robot is immobilised due to wheel slip \cite{ward_classification2007}. This SVM uses wheel-speed and inertial measurements to categorise robot states as ``immobilised,'' ``normal,'' or ``unknown.'' The classification relies on a feature vector that includes the variances of roll rate, pitch rate, and z-axis acceleration, together with the mean wheel angular acceleration. This SVM-based detector can be combined with a model-based slip detector using an Extended Kalman Filter to reduce false-negative detections.

    Beyond supervised classification, unsupervised learning methods, such as Self-Organising Maps, K-means clustering, and autoencoders, can also be used to categorise slip \cite{Kruger2019}. These techniques use measurements from \glspl{imu}, wheel encoders, and motor currents without requiring labelled slip classes. However, the resulting clusters must still be interpreted and associated with physically meaningful slip conditions.

    Another data-driven method for slip compensation uses visual-inertial-wheel odometry \cite{Reginald2025}. This approach applies Gaussian-process regression and Long Short-Term Memory networks to compensate for errors in wheel odometry. The compensated wheel odometry is then fused with inertial and visual odometry using a multi-state constraint Kalman filter. A feature-confidence estimator is also used to reduce the influence of unreliable visual measurements.

    Gaussian-process-based learning can also be used to model wheel-odometry and \gls{imu} errors \cite{Brossard2019}. An Extended Kalman Filter then integrates the output of the learned model with the inertial measurements. Unlike the previous learning-based methods, this approach does not explicitly detect or estimate slip. Instead, slip is treated indirectly as one possible source of wheel-odometry error.

    Overall, learning-based approaches can represent complex slip and odometry errors without requiring a complete analytical model of the wheel-terrain interaction. Supervised approaches require labelled data, while unsupervised approaches avoid explicit labels but still require representative sensor data and interpretation of the resulting clusters. The learned models may also depend on the robot platform, sensor configuration, and terrain conditions used during data collection. Furthermore, the reviewed methods mainly classify complete-wheel slip or learn an aggregate odometry correction and do not explicitly estimate slip at individual rimless-wheel spokes.
    
\section{Research Gap}
\label{sec:sota_gap}

    The reviewed \textbf{wheel--inertial odometry methods} mainly estimate robot motion using conventional wheels with continuous wheel--ground contact. These methods either treat wheel odometry as a single measurement, estimate slip using a wheel or terrain model, or reduce the influence of inconsistent wheel measurements. However, a \gls{rw} interacts with the terrain through separate spokes, and the active contact changes as the wheel rotates. Therefore, the motion of a rimless-wheeled robot depends on the location and velocity of the active contact spoke and cannot be represented directly using the continuous-contact assumptions of conventional wheel models \cite{Das2021, Liu2020, Ward2008, Gargiulo2021}.

    \textbf{Contact-aided state-estimation methods} use foot or wheel contact information to reduce inertial drift. Methods developed for legged robots generally assume that a detected foot remains stationary, while methods developed for wheeled-legged robots model a contact point that moves continuously along the wheel surface. These contact models do not directly address the selection of an active contact from multiple possible rimless-wheel spokes. In addition, existing slip-aware methods mainly estimate or reject slip at the level of a complete wheel, a wheel set, or a foot contact. They do not evaluate the consistency of the encoder-derived velocity of each potential rimless-wheel contact point or reject an individual spoke measurement when it is inconsistent with the predicted robot motion \cite{Hartley2018ContactAidedInEKF, Li2023, Zhou2025, Yoon2024, Kim2025}.

    \textbf{Model-based slip-estimation methods} require assumptions about the wheel, vehicle, terrain, or wheel--terrain interaction. Learning-based methods reduce the need for an analytical slip model, but they require representative training data and may depend on the robot platform and terrain conditions used during data collection. A research gap therefore remains for a purely proprioceptive state-estimation method that accounts for the intermittent contact geometry of \glspl{rw}, identifies the active contact spoke, and reduces the influence of slip-affected contact measurements without requiring a detailed terrain model, labelled training data, or external position measurements during estimation \cite{Kruger2019, Reginald2025, Brossard2019}.

    This work addresses this gap by combining \gls{imu}-based state propagation in an \gls{inekf} with \gls{pcv} measurements derived from the wheel encoders and the rimless-wheel spoke geometry. Candidate contact spokes are determined from the wheel configuration, and the consistency between the measured and predicted contact velocities is evaluated during the filter correction step. The \gls{md} of the contact-velocity innovation is used to identify and reject measurements that may be affected by slip or incorrect contact-spoke selection. This provides a contact-aware state-estimation approach at the individual spoke level using only onboard proprioceptive measurements. The proposed approach and its implementation for the Coyote~3 \gls{rwr} are presented in the following chapters.

\section{Problem Statement}

The baseline odometry estimates the rover motion using wheel-encoder and \gls{imu} measurements. However, it does not explicitly account for wheel slip or incorrect contact-spoke selection. When a spoke slips, the wheel rotation can be interpreted as rover motion even though the corresponding displacement has not occurred. This can introduce errors in the estimated position and travel distance.

Most wheel-inertial and slip-aware odometry methods in the literature are developed for conventional wheels with continuous ground contact. Contact-aided methods are mainly developed for legged robots with stationary foot contacts or for wheeled-legged robots with a contact point that moves continuously along the wheel surface. These assumptions do not directly represent rimless wheels, where the ground contact changes between individual spokes \cite{Hartley2018ContactAidedInEKF,Liu2020,Zhou2025}. Based on these limitations, a \gls{pcv}-based odometry approach using the \gls{inekf} framework is developed. The rover state is propagated using \gls{imu} measurements, then the wheel velocity and spoke geometry are used to calculate the measured \gls{pcv}. The predicted and measured \gls{pcv} are compared, and their difference is calculated as innovation. The \gls{md} of the covariance of this innovation is used to reject measurements that may be affected by slip or incorrect contact-spoke selection.

The proposed contact-velocity odometry is compared with the baseline odometry using experimental datasets collected across different terrains. The evaluation is used to determine whether the proposed sensor-fusion approach improves the odometry estimation and to identify its limitations.
    
\end{document}
