%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}
    \chapter{Methodology} \label{chap:methodology}
    
    This chapter provides details on the proposed approach to the odometry estimation problem mentioned in the problem statement chapter \ref{chap:problem_statement}. The details of \gls{inekf}-based Odometry, the proposed measurement update step for the \gls{inekf}, the Mahalanobis-distance-based measurement rejection, the contact spoke estimation, and the limitations of the proposed approach are presented in this chapter.
    
    \section{Invariant EKF based Odometry}
    
    The proposed approach for odometry estimation in \gls{rwr} uses the proprioceptive sensors available on the rover. The \gls{imu} and the wheel encoders are used in the fusion process. Kalman filters operate in two steps: the propagation step and the correction step. In this thesis, we use the \gls{imu} data for the propagation step and the wheel encoder data is used to construct the measurement used in the correction step. The \gls{inekf} estimates the pose of the rover's body frame with respect to the world frame, the velocity of the rover and the \gls{imu} biases. The filter's state vector and the details on the propagation step are provided in the following subsections.
    
    \todo[inline]{The invariance part needs to be explained, and the explanation related to Lie group and their advantage should also be explained with reference }
    
    \subsection{Filter State Vector}    
    The state vector $X$ contains the quantities that need to be estimated by the filter. In this work, the state vector contains position, orientation, velocity and \gls{imu} sensor biases. These quantities are required to estimate the rover motion. The filter state is constructed as shown in the Equation~\ref{eq:stateVector}. 
    
    \begin{equation}
        X = [R, V, P, b_a, b_g]
        \label{eq:stateVector}
    \end{equation}
    where, $R$ is the rotation matrix, that represents the orientation of the body frame with respect to the world frame. $V$ is the velocity vector of the body frame in the world frame, $P$ is the position vector of the body frame in the world frame, $b_a$ is the accelerometer bias, $b_g$ is the gyroscope bias. 
    
    \subsection{Propagation step}
    
    The propagation step propagates the current filter state defined in Equation~\ref{eq:stateVector} to the next time step using the state transition matrix and the process noise matrix. The elements of the state vector are propagated through time using the \gls{imu} data. This thesis does not modify the propagation step and uses the original implementation \cite{REPO-inekf-github}. The approach presented by Hartley et al. \cite{Hartley2018ContactAidedInEKF}, the authors of the original \gls{inekf} implementation, augment the state by adding the position of contact points to the state vector. In contrast, this thesis keeps the state vector unchanged and only estimate the state variables mentioned in Equation~\ref{eq:stateVector}.
    
    \section{Measurement Update Step}
    
    This section presents the core methodology of the proposed approach. For the fusion process, the \gls{pcv} vector is calculated. This is the velocity of a wheel spoke tip expressed in the body frame in a \gls{rwr} system. The \gls{pcv} can also be referred to as spoke-tip velocity or contact velocity. The term "Potential" comes from the possibility that any one of the spokes in a wheel can be in contact depending on the terrain. The \gls{pcv} is calculated in two forms. The \gls{pcv} vector, calculated from the filter state vector, after the prediction step of the \gls{inekf} is called predicted \gls{pcv}. In the update step, the \gls{pcv} vector is calculated from the wheel encoder measurements called measured \gls{pcv}. Within the \gls{inekf}, the predicted and measured \gls{pcv} vectors are used to calculate the residual, also known as the innovation.
    
    \subsection{Predicted Potential Contact Velocity} \label{subsec:predictedPCV}
    
    The predicted \gls{pcv} is calculated after the propagation step of the \gls{inekf}. The vector quantities used in the predicted \gls{pcv} calculation are expressed in the body frame of the rover. This is done to ensure the consistency between predicted and measured \gls{pcv}.
    The velocity of a \gls{pc} fixed to the rover can be calculated with Equation~\ref{eq:velocityOfaPointInRigidBody}. This equation is not used to calculate the predicted \gls{pcv}. The negated form of this equation is used to calculate the predicted \gls{pcv} as shown in Equation~\ref{eq:measurementFunction}.
        
    \begin{equation}
        v_{pc,rigid}^{body} = ([R_{body}^{world}]^T  v_{body}^{world} + (\omega_{body}^{body} \times r_{pc}^{body}))
        \label{eq:velocityOfaPointInRigidBody}
    \end{equation}
    
    \begin{equation}
        v_{pc, predicted}^{body} = -([R_{body}^{world}]^T  v_{body}^{world} + (\omega_{body}^{body} \times r_{pc}^{body}))
        \label{eq:measurementFunction}
    \end{equation}
    
    where $r_{pc}^{body}$, is the position vector of the \gls{pc} from the body frame of the rover expressed in the body frame, $\omega_{body}^{body}$ is the angular velocity of the body expressed in the body frame. $v_{body}^{world}$ represents the linear velocity of the body frame expressed in the world frame. $R_{body}^{world}$ is the orientation of the body frame expressed in the world frame and $[R_{body}^{world}]^T$ transforms the $v_{body}^{world}$ to the $v_{body}^{body}$. The $R_{body}^{world}$ and the $v_{body}^{world}$ are part of the filter's state vector \ref{eq:stateVector}.
    
    The $v_{pc, rigid}^{body}$ is the velocity of a point on the rover moving through the world expressed in the body frame. This quantity is calculated in Equation~\ref{eq:velocityOfaPointInRigidBody}. The negation of this equation represents the velocity with which the world is moving under the rover, also referred to as the measurement function $h(x)$. This quantity is calculated using the Equation~\ref{eq:measurementFunction}. 
    
    To understand this negation, we should look at an example. Imagine we have rover that is moving forward with a velocity of $+2 m/s$, for the sake of simplicity, our imaginary rover does not have rotational velocity. According to rigid body kinematics, the velocity of a point attached to the rover will also be $+2 m/s$. From the rover's body frame of reference, the forward motion of the body will be seen as the ground moving backwards under the rover. Mathematically, this can be represented by the negation of Equation~\ref{eq:velocityOfaPointInRigidBody}, resulting in the velocity of the ground moving under the point of contact to be $-2 m/s$. This shift in perspective is done to compare the predicted and the measured \gls{pcv}
        
    
    \subsection{Measured Potential Contact Velocity} \label{subsec:measuredPCV}
    
    The measured \gls{pcv} is calculated using the wheel's angular velocity expressed as a vector and the position vector of \gls{pc}. The position vector is calculated from the wheel center to the tip of the spoke. The measured \gls{pcv} vector is perpendicular to the position vector and points in the direction of rotation of the wheel. The calculation of the measured \gls{pcv} is shown in Equation~\ref{eq:MeasuredPCVCalculation}. The measured \gls{pcv} vector is transformed to the body frame to be consistent with the predicted \gls{pcv} and before the residual calculation.
    
    \begin{equation}
        v_{pc, measured}^{body} = R_{wheel}^{body} (\omega_{wheel}^{wheel} \times {}^{}r_{pc}^{wheel})
        \label{eq:MeasuredPCVCalculation}
    \end{equation}
    
    From, the rover's body frame of reference, $v_{pc, measured}^{body}$ represents the velocity of the ground moving under the point of contact. This is the same quantity as the $v_{pc, predicted}^{body}$, so the difference between them can be calculated as the residual.
      
    \subsection{Residual Calculation} \label{subsec:residual}
    
    The residual, also referred to as the innovation is calculated as the difference between the measured and the predicted \gls{pcv} when both quantities are expressed in the body frame of reference. The residual is used in the \gls{inekf} to update the state vector. The implementation details are provided in Chapter~\ref{chap:implementation}.
    
    \begin{equation}
        r_{pcv} =  {v}_{pc, measured}^{body} - {v}_{pc, predicted}^{body} 
    \label{eq:pcvResidual}
    \end{equation}
    
    If the $r_{pcv}$ is small, the wheel measurements are consistent with the filter's predicted rover motion. If the residual is large, then the wheel measurements are inconsistent with the filter's prediction. This could mean wheel measurements may be affected by errors such as wheel slip, an inconsistent contact assumption or noise in wheel measurements. The residual is also used in the \gls{md} calculation which is later used in slip rejection and contact estimation.

    \section{Measurement Model and Jacobian}
        
    The measurement model, also referred to as the measurement function $h(x)$, is given by Equation~\ref{eq:measurementFunction}. The $h(x)$ corresponds to the predicted \gls{pcv} which is used by \gls{inekf} to calculate the predicted measurement from its state vector that can be compared to a measured \gls{pcv} derived from the wheel measurements. The $h(x)$ is also used to derive the measurement Jacobian. For better readability, the measurement function is rewritten in Equation~\ref{eq:measurementFunction-2}.
    
    The measurement Jacobian $H$ is the partial derivative of the measurement function $h(x)$ with respect to the error state $\xi$ variables evaluated at the current state $\hat{X}$ as shown in Equation~\ref{eq:MeasurementJacobian-1}. The measurement Jacobian linearizes the measurement function at the current state. The measurement Jacobian describes the relationship between a small perturbation in the state vector and the predicted \gls{pcv} value. The measurement Jacobian calculation and its results for this measurement function is shown in equations \ref{eq:MeasurementJacobian-2} and \ref{eq:MeasurementJacobian-3} respectively.
        
        \begin{equation}
            h(x) = -([R_{body}^{world}]^T  v_{body}^{world} + (\omega_{body}^{body} \times r_{pc}^{body}))
            \label{eq:measurementFunction-2}
        \end{equation}
    
        \begin{equation}
            \xi = \begin{bmatrix}
            \delta R \\
            \delta V \\
            \delta P \\
            \delta acc\_bias \\
            \delta gyro\_bias
            \end{bmatrix}
            \label{eq:ErrorStateVector}
        \end{equation}
        
        \begin{equation}
            H = \frac{\partial h(x)}{\partial \xi}\bigg|_{\hat{X}}
            \label{eq:MeasurementJacobian-1}
        \end{equation}
        
        where, $\xi$ is the left invariant error state, refer to Section~\ref{sec:LeftAndRightInvariantError} for details, $\hat{X}$ is the current state estimate, and $h(x)$ is the measurement function. The evaluation of Equation~\ref{eq:MeasurementJacobian-2} to get the result in Equation~\ref{eq:MeasurementJacobian-3} is discussed in detail in the appendix \ref{appendix:MeasurementJacobianDerivation}.
        
        \begin{equation}
            H = \begin{bmatrix}
            \frac{\partial h(x)}{\partial \delta R} & \frac{\partial h(x)}{\partial \delta V} & \frac{\partial h(x)}{\partial \delta P} & \frac{\partial h(x)}{\partial \delta acc\_bias} & \frac{\partial h(x)}{\partial \delta gyro\_bias}
            \end{bmatrix}
            \label{eq:MeasurementJacobian-2}
        \end{equation}
        
        \begin{equation}
            H = \begin{bmatrix}
                -[R_{world}^{body} * v_{body}^{world}]_\times & -\mathbf{I} & \mathbf{0} & \mathbf{0} & -[r_{pc}^{body}]_\times
            \end{bmatrix}
            \label{eq:MeasurementJacobian-3}      
        \end{equation}
        
        The validity of the measurement Jacobian is verified by comparing the analytical Jacobian with a numerical Jacobian. The numerical Jacobian is calculated by perturbing the state vector and calculating the change in the measurement function. The analytical and numerical Jacobians are found to be consistent with each other. The numerical jacobian calculation is provided in the appendix \ref{appendix:NumericalJacobianCalculation}. 
        \todo[inline]{Provide the numerical jacobian calculation in the appendix}
       
        \section{Slip-aware Measurement Rejection}
        
        The measured \gls{pcv}, calculated from the wheel encoder measurements, can be affected by wheel slip. This wheel slip can occur due to several factors such as slippery terrain, compliant terrain like soft sand, loose and uneven terrain like pebbles, inclined terrain, very high wheel speeds or a combination of these factors. The nature of slip and the details required to understand the wheel slip is discussed in Chapter~\ref{chap:background}. In the \gls{inekf} framework, the slip-induced inconsistencies may be identified when the measured \gls{pcv} vector calculated from the wheel encoders is inconsistent with the predicted \gls{pcv} computed from the filter state. 
        
        This inconsistency can be calculated using the \gls{md}. The \gls{md} measures residual between a measurement and its predicted values weighted by the inverse residual covariance \cite{SITE-mathworks-mahalanobis}. The residual calculated in Section~\ref{subsec:residual} is used to calculate the \gls{md}. When the squared \gls{md} is higher than a threshold value (details are provided in the Chapter~\ref{chap:implementation})\todo[inline]{Include the chi2 threshold in the implementation chapter} then the measured \gls{pcv} is rejected from the update step, thereby preventing the inconsistent wheel encoder measurements from a slip event to be used \cite{chang2014MahalanobisFormula}.
        
        \begin{equation}
        S = HPH^T + R
        \label{eq:innovationCovariance}
        \end{equation}
        
        \begin{equation}
        d_M^2 = r^T S^{-1} r
        \label{eq:mahalanobisDistanceSquared}
        \end{equation}
        
        where $S$ is the residual or innovation covariance, $H$ is the measurement Jacobian calculated in the previous section, $P$ is the state covariance estimated by the filter, $R$ is the measurement noise covariance, and $r$ is the residual.
        
        \section{Contact estimation}
        
        The design of rimless wheel, being a hybrid between wheeled and legged systems have the swing phase and the contact phase. Estimating which spoke is currently is in contact for a given wheel is a vital information for the Contact Velocity Odometry estimation. There is a traditional way of estimating the contact spoke by using the assumption that the spoke that is close to the ground is in contact. This can be calculated by calculating the robot kinematics, using the joint positions of the wheel. The pose of each spoke with respect to the body frame can be calculated and the spoke in contact can be estimated. This assumption works well in a flat, solid terrain but on an inclined or rough terrain with obstacles, this assumption does not hold. 
        
        The use of Kalman filter for odometry estimation provides an opportunity to use its framework to estimate the feet that are in contact.
        In order to understand the contact estimation process, one must understand the Kalman filter framework implemented in this thesis and explained in \ref{subsec:inekf_implementation}\todo[inline]{Write a subsection on InEKF implementation}. The Kalman filter calculates the term innovation, in the case of this thesis, the innovation is calculated as the difference between the velocity of spoke from the filter's prediction and the velocity of the spoke 
    
    
    
\end{document}
