%!TEX root = ../report.tex
\documentclass[report.tex]{subfiles}
\begin{document}
    \chapter{Solution}

    \section{Proposed algorithm}
        
        The proposed approach modifies the existing InEKF library by adding a new update method that uses Potential Contact Velocity vector as a measurement to correct the filter state. As discussed in the background chapter, the PCV vector is calculated using the wheel encoder measurements and compared against the PCV calculated from the filter state. The difference between these two sources of information is used to get the innovation/residual for the update/correction step of the filter. 

    \section{Implementation details}
    
    \subsection{Filter update step implementation}
    
        This section describes the implementation details related to the update step of the \gls{inekf}. The update step, as discussed in the methodology chapter \ref{chap:methodology}, utilizes the \gls{pcv} as the measurement. Firstly, we use the state vector of the \gls{inekf} to calculate the expected \gls{pcv}, and we use the wheel velocities(Rotational speed of the wheels) to calculate the measured \gls{pcv}. The following subsection will discuss the design and implementation choices used in the above-mentioned calculations.
    
    \subsection{Expected Potential Contact Velocity Calculation}
        
        The filter state consists of $[R, V, P, acc\_bias, gyro\_bias]$, the expected \gls{pcv} need to be calculated with this information. All the calculations we do related to the \gls{inekf} are  body centric calculations. In order to calculate the velocity of a point in a rigid body relative to the body frame, we need to represent the velocity of body frame at the point we need the velocity at. In our case, our robot model has a body frame, which is located at the center of the rover, and we need to calculate the velocity of the \gls{pc} relative to the body frame. 
        
        This calculation can be done step by step by using equation \ref{eq:ExpectedPCVCalculation}. This operation can also be done using \gls{kdl} using Twists \ref{subsec:twists} by utilizing the adjoint representation of the transformation matrix using the equation \ref{eq:TwistTransformationEquation}. The implementation version of the twist transformation is provided in \ref{eq:ExpectedPCVCalculationTwist}
    
        \begin{equation}
            v_{pc}^{body} = v_{body}^{body} + (\omega_{body}^{body} \times {}^{}r_{pc}^{body})
            \label{eq:ExpectedPCVCalculation}
        \end{equation}
                
        \begin{equation}
            \begin{bmatrix}
            \boldsymbol{\omega}_{pc}^{body} \\
            \mathbf{v}_{pc}^{body}
            \end{bmatrix}
            =
            \begin{bmatrix}
            I & 0 \\
            \left[\,\mathbf{p}_{pc}^{body}\,\right]_\times & I
            \end{bmatrix}
            \begin{bmatrix}
            \boldsymbol{\omega}_{body}^{body} \\
            \mathbf{v}_{body}^{body}
            \end{bmatrix}
            \label{eq:ExpectedPCVCalculationTwist}            
        \end{equation}
    
    \subsection{Measured Potential Contact Velocity Calculation}
        
    \subsection{Fixed Wheel Frames}
    
    The fixed wheel frames are added in the robot URDF model at the location of motors. This fixed wheel frames have the same pose as the initial pose of the rotating wheel frame. The fixed wheel frames are added to get the position vector to the spoke of the wheel from the center of the wheel. These position vectors are later used to calculate the measured \gls{pcv}. This is by performing a cross product of wheel's angular velocity vector (the vector form of rotation speed) and the position vector of \gls{pc}. The \gls{pcv} vector will be perpendicular to the spoke and point in the direction of rotation of the wheel. This is the only purpose of the fixed wheel frames. The calculation of the measured \gls{pcv} is shown in equation \ref{eq:MeasuredPCVCalculation}.
    
    \begin{equation}
        v_{pc}^{wheel} = \omega_{wheel}^{wheel} \times {}^{}r_{pc}^{wheel}
        \label{eq:MeasuredPCVCalculation}
    \end{equation}
    
    
    \subsection{Measurement Jacobian}
    
        The measurement jacobian $H$ is the partial derivative of the measurement function $h$ with respect to the error state $\xi$ variables at the state $\hat{X}$ as shown in equation \ref{eq:MeasurementJacobian-1}. The measurement jacobian linearizes the measurement function at the current state. The measurement function $h(x)$ is given by the equation \ref{eq:MeasurementFunction}. Measurement jacobian for this measurement function is obtained by taking the partial derivative of the measurement function with respect to the error state variables \ref{eq:ErrorStateVector}. The measurement jacobian calculation and its results for this measurement function is shown in equations \ref{eq:MeasurementJacobian-2} and \ref{eq:MeasurementJacobian-3} respectively.
    
        \begin{equation}
            h(x) = - (v_{body}^{body} + (\omega_{body}^{body} \times {}^{}r_{pc}^{body}))
            \label{eq:MeasurementFunction}
        \end{equation}
        
        \begin{equation}
            \xi = \begin{bmatrix}
            \delta R \\
            \delta V \\
            \delta P \\
            \delta acc\_bias \\
            \delta gyro\_bias
            \end{bmatrix}
            \label{eq:ErrorStateVector}
        \end{equation}
        
        \begin{equation}
            H = \frac{\partial h}{\partial \xi}\bigg|_{\hat{X}}
            \label{eq:MeasurementJacobian-1}
        \end{equation}
        where, $\xi$ is the left invariant error state, $\hat{X}$ is the current state estimate, and $h$ is the measurement function. 
        
        
        \begin{equation}
            H = \begin{bmatrix}
            \frac{\partial h}{\partial \delta R} & \frac{\partial h}{\partial \delta V} & \frac{\partial h}{\partial \delta P} & \frac{\partial h}{\partial \delta acc\_bias} & \frac{\partial h}{\partial \delta gyro\_bias}
            \end{bmatrix}
            \label{eq:MeasurementJacobian-2}
        \end{equation}
        
        \begin{equation}
            H = \begin{bmatrix}
                \mathbf{0} & -\mathbf{I} & \mathbf{0} & \mathbf{0} & -[r_{pc}^{body}]_\times
            \end{bmatrix}
            \label{eq:MeasurementJacobian-3}      
        \end{equation}
       
        The evaluation of equation \ref{eq:MeasurementJacobian-2} to get the result in equation \ref{eq:MeasurementJacobian-3} is discussed in detail in the appendix \ref{appendix:MeasurementJacobianDerivation}.

    \subsection{Slip elimination}
    
    The slippage is discussed in detail in the background chapter \ref{subsec:backgroundSlippage}

    
\end{document}
